### 2024-11-30 Final/Baseline_Hockey.Rmd
Baseline Model

---
title: "R.Final_PenalizedRegression"
author: "Connor Logue"
date: "2024-11-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data
```{r}
data = read.csv("data.csv", header = TRUE)
data
```

### EDA  & Cleaning
```{r}
paste("Number of Instances:", nrow(data))
paste("Number of Variables:", ncol(data))
```

```{r}
head (data)
```

### Baseline
```{r}
base <- lm(Salary_log ~ FO_pct + G + A + GF + GA + plus_mins + TOI + iSF + Yrs_in_NHL + Ht + Wt + Grit + iHF + iHA + iBLK + PIM + GP, data = data)
```

``` {r}
summary(base)
```

```{r}
vif(base)
mean(vif(base))
```

```{r}
RSS <- c(crossprod(base$residuals))
MSE <- RSS / length(base$residuals)
MSE
```

```{r}
base2 <- lm(Salary_log ~ G + A + GF + GA + plus_mins + TOI + iSF + Yrs_in_NHL + GP, data = data)
summary(base2)
```
```{r}
vif(base2)
mean(vif(base2))
```