### 2024-12-01 Final/Baseline_Penalized.Rmd
Bedrock, Elastic Model Selection, Baseline, Ridge Shrinkage

---
title: "Con"
author: "Connor Logue"
date: "2024-12-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Libraries
```{r}
library(dplyr)
library(glmnet)
library(car)
```

### Data
```{r}
data = read.csv("combined.csv", header = TRUE)
data
```

### Cleansing
```{r}
data$Yrs_in_NHL <- 2016 - data$DftYr
data$A = data$A1 + data$A2

data$GF[is.na(data$GF)] <- median(data$GF, na.rm=TRUE)
data$GA[is.na(data$GA)] <- median(data$GA, na.rm=TRUE)
data$iSF[is.na(data$iSF)] <- median(data$iSF, na.rm=TRUE)
data$Yrs_in_NHL[is.na(data$Yrs_in_NHL)] <- 0
data$iHA[is.na(data$iHA)] <- median(data$iHA, na.rm=TRUE)

data <- data[data$Yrs_in_NHL > 3,]
data <- data[data$Is_Offense == 1,]

scale_numeric <- function(x) x %>% mutate_if(is.numeric, function(y) as.vector(scale(y)))
scaled_data <- data %>% scale_numeric()
```

```{r}
bedrock <- lm(Salary_log ~ FO_pct + G + A + GF + GA + plus_mins + TOI + iSF + Yrs_in_NHL + Ht + Wt + Grit + iHF + iHA + iBLK + PIM + GP, data = scaled_data)
summary(bedrock)

e     <- resid(bedrock)
Yhat  <- predict(bedrock)
s     <- summary(bedrock)$sigma
es    <- e/s
```

```{r}
plot(Yhat, e, type = 'n', main = 'Fitted vs Residual')
grid()
abline(h = 0, lwd = 2, col = 'gray')
points(Yhat, e, pch = 16, col = 'darkblue')
```

```{r}
qq_model  <- qqnorm(es, plot.it = FALSE)
qqplot(qq_model$x, qq_model$y, xlab = 'E(e*)', ylab = 'e*', pch = 16, col = 'darkblue',
       main = 'Q-Q Plot')
qqline(qq_model$y, col = 'gray', lwd = 2)
grid()
```

```{r}
vif(bedrock)
mean(vif(bedrock))
```
```{r, fig.width=10, fig.height=10}
pairwise <- subset(scaled_data, select = c(FO_pct, G, A, GF, GA, plus_mins, TOI, iSF, Yrs_in_NHL, Ht, Wt, Grit, iHF, iHA, iBLK, PIM, GP))

library(corrplot)
M = cor(pairwise)
corrplot(M, method = 'number')
```

### Penalized Regression
``` {r}
library(glmnet)
X         <- model.matrix(Salary_log ~ FO_pct + G + A + GF + GA + plus_mins + TOI + iSF + Yrs_in_NHL + Ht + Wt + Grit + iHF + iHA + iBLK + PIM + GP -1, data = scaled_data)
Y         <- scaled_data$Salary_log

ridge_cv   <- cv.glmnet(x = X, y = Y, family = "gaussian", alpha = 0, type.measure = "mse", nfolds = nrow(scaled_data), grouped = FALSE)
elastic_cv <- cv.glmnet(x = X, y = Y, family = "gaussian", alpha = 0.5, type.measure = "mse", nfolds = nrow(scaled_data), grouped = FALSE)
lasso_like_cv <- cv.glmnet(x = X, y = Y, family = "gaussian", alpha = 0.9, type.measure = "mse", nfolds = nrow(scaled_data), grouped = FALSE)
lasso_cv <- cv.glmnet(x = X, y = Y, family = "gaussian", alpha = 1, type.measure = "mse", nfolds = nrow(scaled_data), grouped = FALSE)

results <- matrix(c(ridge_cv$lambda[which.min(ridge_cv$cvm)], min(ridge_cv$cvm),
              elastic_cv$lambda[which.min(elastic_cv$cvm)],min(elastic_cv$cvm),
              lasso_like_cv$lambda[which.min(lasso_like_cv$cvm)], min(lasso_like_cv$cvm),
              lasso_cv$lambda[which.min(lasso_cv$cvm)], min(lasso_cv$cvm)), 
            nrow = 4, ncol = 2, byrow = TRUE)

rownames(results) = c("Ridge", "Elastic", "Lasso-Like", "Lasso")
colnames(results) = c("Lambda", "MSE")
results
```
```{r}
plot(elastic_cv)
```

```{r}
coef(elastic_cv)
```

```{r}
baseline <- lm(Salary_log ~ G + A + GF + plus_mins + TOI + iSF + Yrs_in_NHL + Wt + GP, data = scaled_data)
summary(baseline)

e     <- resid(baseline)
Yhat  <- predict(baseline)
s     <- summary(baseline)$sigma
es    <- e/s
```

```{r}
plot(Yhat, e, type = 'n', main = 'Fitted vs Residual')
grid()
abline(h = 0, lwd = 2, col = 'gray')
points(Yhat, e, pch = 16, col = 'darkblue')
```

```{r}
qq_model  <- qqnorm(es, plot.it = FALSE)
qqplot(qq_model$x, qq_model$y, xlab = 'E(e*)', ylab = 'e*', pch = 16, col = 'darkblue',
       main = 'Q-Q Plot')
qqline(qq_model$y, col = 'gray', lwd = 2)
grid()
```

```{r}
vif(baseline)
mean(vif(baseline))
```

```{r}
library(glmnet)
X         <- model.matrix(Salary_log ~ G + A + GF + plus_mins + TOI + iSF + Yrs_in_NHL + Wt + GP -1, data = scaled_data)
Y         <- scaled_data$Salary_log

ridge_cv   <- cv.glmnet(x = X, y = Y, family = "gaussian", alpha = 0, type.measure = "mse", nfolds = nrow(scaled_data), grouped = FALSE)

plot(ridge_cv)
```

```{r}
coef(ridge_cv)
```